---
title: "P8105-hw5-zc2556"
author: "Zhe Chen"
date: "2020/11/17"
output: github_document
---

## Problem 1

### Libraries and Basics

```{r, warning=FALSE, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(
  theme_minimal()+
  theme(legend.position = "bottom")
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continous.fill = "viridis"
)

scale_color_discrete = scale_colour_viridis_d()
scale_fill_discrete = scale_fill_viridis_d

```

### Import and Clean the Data Set

```{r}
#read in the dataset
homoicide = 
  read_csv("./homicide-data.csv") %>%
  #redefine the arrest status
  mutate(
    city_state = str_c(city, state, sep = "_"),
    resolved = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    )
  ) %>%
  select(
    city_state,resolved
  ) %>%
  filter(city_state != "Tulsa_AL")
```

Check the data set

```{r}
#give a tidy data frame including the solved/unsolved cases 
aggregate_homo = 
  homoicide %>%
  group_by(city_state) %>%
  summarize(
    hom_total = n(),
    hom_unsolved = sum(resolved == "unsolved")
  )
aggregate_homo
```

Prop test for a single city

```{r}
prop.test(
  aggregate_homo %>% filter(city_state == "Baltimore_MD") %>% pull(hom_unsolved),
  aggregate_homo %>% filter(city_state == "Baltimore_MD") %>% pull(hom_total)
  ) %>%
  broom::tidy() #make it exhibit in one row
  
```

Iterate

```{r}
#using map to give prop tests of each city state
results =
  aggregate_homo %>%
  mutate(
    prop_test = map2(.x = hom_unsolved, .y = hom_total, ~prop.test(x = .x, n = .y)),
    tidy_test = map(.x = prop_test, ~broom::tidy(.x))
  ) %>%
  select(
    -prop_test
  ) %>%
  unnest(tidy_test) %>%
  select(
    city_state, estimate, conf.low, conf.high
  )
results
```

Plot

```{r}
results %>%
  mutate(city_state = fct_reorder(city_state, estimate)) %>%
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point()+
  geom_errorbar(ymin = results$conf.low, ymax = results$conf.high)+
  theme(axis.text = element_text(angle = 90, vjust = 0.5, hjust = 1))

```


## Problem 2

### Import Data Sets with Iteration

```{r, warning= FALSE, message=FALSE}
#store file paths of the control data sets
path_con = 
  str_c("./data/", list.files("data", pattern = "con"))

#store file paths of the experiment data sets 
path_exp = 
  str_c("./data/", list.files("data", pattern = "exp"))

#import both controls and experiments into a single data set
long_data = 
  tibble(
    control = map(path_con, read_csv),
    experiment = map(path_exp, read_csv)
  )

path_con = 
  str_c("./data/", list.files("data", pattern = "con"))

path_exp = 
  str_c("./data/", list.files("data", pattern = "exp"))

long_data = 
  tibble(
    control = map(path_con, read_csv),
    experiment = map(path_exp, read_csv)
  )
```

### Make the Data Tidy

```{r, message=FALSE}
#make it tidy
long_data_tidy = 
  long_data %>%
  mutate(
    subject_id = 1:10
  ) %>%
  pivot_longer(
    control:experiment,
    names_to = "arm"
  ) %>%
  unnest() %>%
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    names_prefix = "week_",
    values_to = "observations"
  )

long_data_tidy
```

### Plot

```{r}
long_data_tidy %>%
  mutate(subject_id = as.factor(subject_id),
         week = as.factor(week)
         ) %>%
  ggplot(aes(x = week, y = observations, group = subject_id, color = subject_id)) +
  geom_path()+
  facet_grid(~arm) +
  labs(
    title = "Observations over 8 Weeks by Treatment Groups",
    x = "Week",
    y = "Observations"
  ) + 
  theme(legend.position = "bottom")
  
```

We made a spaghetti plot to show the observations on each subject over time to compare the effect of the treatment. In control group, we don't observe a strong trend and the change of observations for 8 weeks seems to be random. However, for the experiment group, we can observe a clearly increasing tread, showing the treatment should have some effect compared to the control.   


## Problem 3

```{r}
#set up 
set.seed(621)
n_ent = 30
sigm_ent = 5

#function to find the estimated mean and p-value
sim_mean_p = function(n, mu, sigm) {
  x = rnorm(n, mean = mu, sd = sigm)
  t = broom::tidy(t.test(x))
  fun_res = tibble(
    mu_hat = t$estimate,
    p_value = t$p.value
    )
  return(fun_res)
}

```


5000 for mu = 0

```{r}
#function to stimulate for 5000 times 
stimulate = function(mu) {
  output_mu= vector("list", 5000)
  for(i in 1:5000) {
    output_mu[[i]] = sim_mean_p(n_ent, mu, sigm_ent)
  }
  sim_result = bind_rows(output_mu)
  return(sim_result)
}

#when mu = 0
mu0 = stimulate(0)
mu0
```

Proportion of Rejecting the Null based on 0.05 alpha when mu = 0.

```{r}
#function to find the proportion of rejecting the null
test = function(stimulate_data, true_mu) {
  sim_reject = 
    stimulate_data %>%
      filter(p_value > 0.05) %>%
      summarise(
        prop_reject = n()/5000
      )
  mu =
    tibble(
      true_mu = true_mu
    )
  sim_reject = bind_cols(sim_reject, mu)
  return(sim_reject)
}

#for mu = 0 
test(mu0,0)
```

### Plot the Proportion of Reject VS True Mean

Apply functions 
```{r}
#make a data frame including mu = 1-6
for (i in 1:6) {
  assign(paste0("mu", i), stimulate(i))
}
```

Make a table for plotting
```{r}
#make a data frames including the proportion and true means
mu_all = list(mu1,mu2,mu3,mu4,mu5,mu6)
for (i in 1:6) {
  assign(paste0("prop_table", i), test(as.data.frame(mu_all[i]),i))
}
prop_table = bind_rows(prop_table1, prop_table2, prop_table3, prop_table4, prop_table5, prop_table6)
prop_table
```

Plot

```{r}
prop_table %>%
  ggplot(aes(x = as.factor(true_mu), y = prop_reject, group  = 1)) +
  geom_point() +
  geom_line() +
  labs(
    title = "5000 Times of Stimulation for 6 Means",
    x = "True Mean",
    y = "Proportion of Rejection Times"
  ) + 
  theme(legend.position = "bottom")
```














